---
title: "Models specification"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---


```{r model_spec setup, include = FALSE, eval = TRUE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
source("Scripts/00_setup.R")
```


# Hidden semi-Markov models specifications



In this section, we specify the 3 hidden semi-Markov models (HSMMs) used to decode the app users' time series. We start by specifying the multi-variate model of reproductive events, then simplify this model so that it can be used to decode uni-variate time series of bleeding reports and identify menses from these self-reports, and finally, we specify the model of long-term changes in tracking behavior.


```{r child = '1A_HSMM_R_model_specification.Rmd', cache=TRUE}
```


\newpage

```{r child = '1B_HSMM_M_model_specification.Rmd', cache=TRUE}
```


\newpage


```{r child = '1C_HSMM_T_model_specification.Rmd', cache=TRUE}
```






