---
title: "T model specification"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---


## Model to identify sub-sequences in user's time-series with consistent tracking behavior

The tracking-behavior categories identifying model (`T_hsmm`) is specified to identify long sub-sequences in user's time-series in which they had a consistent tracking behavior. Since we assume, for simplicity of the reproductive event labeling process, that users change behavior around their period, this model decodes time-series of five binary variables. The first one indicates whether a period was reliably identified with the `M_hsmm` and the four other variables indicate whether mucus, temperature, pregnancy tests or LH tests were reported.

The tracking behavior categories, _i.e._ states of this model, are defined based on possible combination of reported variables. In theory, this would lead to $\sum_{k=0}^{4} \binom{4}{k} = 16$ states. However, we simplified to 6 states as some combination of reported variables had similar consequences on the biological states that could or could not be discriminated with these variables. For example, LH tests, temperature and pregnancy tests bring information which can help to discriminate between a cycle with a long follicular phase and a cycle with a short follicular phase but an early pregnancy loss. 

To these 6 states, we add a seventh state which is a transition state. This transition state is a short duration state (approximately the same duration as a period) and transitions from any of the 6 other states are done via this transition state.

```{r T-hsmm}

states = c("b","bp","b_tests","no_temp","no_mucus","full", "transition")
J = length(states)
M = 10*365 # 10 years is the max sojourn

T_hsmm = specify_hsmm(
  J = J,
  state_names = states,
  state_colors = c(
    hsv(0.54, s = 0.6, v = 1),
    hsv(0.54, s = 0.8, v = 1),
    hsv(0.54, s = 0.8, v = 0.9),
    hsv(0.54, s = 0.8, v = 0.8),
    hsv(0.54, s = 0.8, v = 0.7),
    hsv(0.54, s = 0.8, v = 0.6),
    "black"
    ),  #c(rainbow(n = 6, s = 1, v = 0.7), "black"),
  init = c(rep(1, J-1)/(J-1), 0),
  transition = rbind(
    matrix(c(rep(0,J-1),1), nrow = J-1, ncol = J, byrow = TRUE), 
    c(rep(1/(J-1),J-1),0)) %>% set_colnames(states) %>% set_rownames(states),
  sojourn = list(type = "ksmoothed_nonparametric", 
                 d = cbind(
                   rbind(matrix(0, nrow = 9*32, ncol = J-1), matrix(1/M, nrow = M, ncol = J-1)),
                   matrix(c(rep(1/5,5),rep(0,9*32+M-5)), nrow = 9*32 + M, ncol = 1))),
  marg_em_probs = list(
                                                                              #b      bp      b_test  no_temp no_mucus full   transition 
    any.menses = list(type = "binom", params = list(size = rep(1,J), prob = c(1/5,    1/5,    1/5,    1/5,    1/5,     1/5,     1    ))),
    any.preg   = list(type = "binom", params = list(size = rep(1,J), prob = c(0,      1/100,  1/100,  1/100,  1/100,   1/100,   1/100))),
    any.LH     = list(type = "binom", params = list(size = rep(1,J), prob = c(0,      1/1000, 1/60,   1/100,  1/100,   1/100,   1/100))),
    any.mucus  = list(type = "binom", params = list(size = rep(1,J), prob = c(0,      1/1000, 1/1000, 1/100,  1/1000,  1/10,    1/50 ))),
    any.temp   = list(type = "binom", params = list(size = rep(1,J), prob = c(0,      1/2000, 1/2000, 1/2000, 1/100,   2/10,    1/20 )))
  )
)

save(T_hsmm, file = "../Data/models/T_hsmm.Rdata")

```

```{r T-hsmm-layout, echo = FALSE}

l = rbind(
  c(cos(5*pi/6), sin(5*pi/6)),
  c(cos(1*pi/2), sin(1*pi/2)),
  c(cos(1*pi/6), sin(1*pi/6)),
  c(cos(-1*pi/6), sin(-1*pi/6)),
  c(cos(-1*pi/2), sin(-1*pi/2)),
  c(cos(-5*pi/6), sin(-5*pi/6)),
  c(0,0)
)

```



```{r T-hsmm-graph-model, fig.width=10, fig.height=10, out.width='50%', fig.cap="Graph of the T-hsmm model", echo = FALSE}

plot_hsmm_transitions(model = T_hsmm, label_color = "white")

```



```{r T-hsmm-sim-viz, fig.width=12, fig.height=3, fig.cap="Simulated sequence with the T-hsmm."}

Ti_sim = simulate_hsmm(model = T_hsmm, n_state_transitions = 7, seed = 2)
plot_hsmm_seq(X = Ti_sim, model = T_hsmm, add_state_color_legend = TRUE)

```


```{r T-hsmm-sim-viz2, echo = FALSE, fig.width=12, fig.height=3, fig.cap="Zoom on a state transition in a sequence simulated with the T-hsmm", eval = FALSE}

plot_hsmm_seq(X = Ti_sim %>%  filter(t %in% 1000:2000), model = T_hsmm, add_state_color_legend = TRUE)

```

