---
title: "Datasets"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---


```{r datasets-setup, include = FALSE, eval = TRUE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
source("Scripts/00_setup.R")
```

# Datasets


## Synthetic data generation


To generate synthetic data, we first define a number of users and some characteristics defining these fictive users such as their highest tracking frequency, controlled with a parameter $\alpha$, and menstrual characteristics such as the length and regularity of their follicular and luteal phases or the noise in their temperature measurements.

Because we want to evaluate the performance of our framework given different level of missing data in a realistic context, we define the effect of the parameter $\alpha$ on the missing probabilities as follow:


$$ q^*_{j,k} = q_{j,k} - \beta \ q_{j,k} + \beta \ \delta  $$

where

$q_{j,k}$ are realistic (expected in real-life data, see fig \@ref(fig:datasets-expected-missingness)) estimates of the missing probabilities with $q_{j,k} = P(X_k \text{ is missing}|S = j)$, $\beta = \| \frac{3^{\alpha} - 1}{3^{\alpha}+1} \|$ and $\delta = 0$ if $\alpha < 0$ and 1 otherwise.

The parameter $\alpha$ takes values in $\mathbb{R}$. When $\alpha$ is zero, the missing probabilities are identical to the expected missing probabilities from a typical app users whose purpose is to identify their fertile window. When $\alpha$ is negative, the missing probabilities are smaller than the expected ones, when alpha is positive they are bigger. In the extreme cases, a value of $-\infty$ means that the variables are never missing and a value of $+\infty$ means that the variables are always missing. When $\alpha = -1$, $q^* = \frac{1}{2} q$: the variables are missing only about half the time they were expected to be missing. When $\alpha = 1$, variables are missing with a probability of at least 50%.


__Expected missingness__

```{r datasets-expected-missingness, echo = FALSE, results='hide', fig.height=3, fig.cap="Missing probabilities for each variable and state expected for a typical users whose tracking purpose is to identify their fertile window and detect pregnancies early on."}

load("../Data/models/R_hsmm.Rdata", verbose = TRUE)

pj_mat = matrix(R_hsmm$censoring_probs$p, 
                nrow = nrow(R_hsmm$censoring_probs$q), 
                ncol = R_hsmm$J, 
                byrow = TRUE)
qjk = pj_mat + (1-pj_mat) * R_hsmm$censoring_probs$q

qjk_long = qjk %>% as.data.frame() %>% 
  mutate(variable = rownames(qjk)) %>% 
  pivot_longer(., -variable, names_to = "state_name", values_to = "q" ) %>% 
  mutate(state_name = state_name %>% factor(., levels = R_hsmm$state_names), 
         state = state_name %>% as.numeric(),
         state_color = R_hsmm$state_colors[state])

ggplot(qjk_long, aes(x = state_name, y = q, fill = state_color)) + geom_bar(stat = "identity") + facet_grid(variable ~ .) + ylim(c(0,1)) + scale_fill_identity()

```

__Synthetic users characteristics__

```{r datasets-synthetic-users}

N_per_alpha = 30 
alpha = rep(c(-Inf, -1, -0.5, 0, 0.5, 1, 2, 4), N_per_alpha)
N_users = length(alpha)

s_users = data.frame(
  seq_id = 1:N_users, 
  alpha = alpha,
  temp_sd = runif(N_users, min = 0.05, max = 0.3),
  mean_lE_sojourn = rnorm(N_users, mean = 10, sd = 3),
  sd_lE_sojourn = 0.2 + rpois(N_users, lambda = 2),
  mean_Lut_sojourn = rnorm(N_users, mean = 11, sd = 1),
  sd_Lut_sojourn = 0.2 + rpois(N_users, lambda = 0.75)
  )

rm(alpha)

```

__Generating time-series__


```{r datasets-Xsim, warning=FALSE}

# 139 sec for 64 time-series
Xsim = purrr::map_dfr(
  .x = s_users$seq_id,
  .f = function(sid){
    #cat("\n",sid, "\n")
    modified_hsmm = R_hsmm
    
    # modifying censoring probabilities
    alpha = s_users$alpha[s_users$seq_id == sid]
    beta = abs((3^alpha - 1)/(3^alpha + 1))
    delta = ifelse(alpha > 0, 1, 0)
    modified_hsmm$censoring_probs$p = 0 *modified_hsmm$censoring_probs$p
    modified_hsmm$censoring_probs$q = qjk * (1 - beta) + beta * delta
    
    # modifying emission distributions
    modified_hsmm$marg_em_probs$temp$params$sd = rep(s_users$temp_sd,R_hsmm$J)
    # modifying sojourns
    M = length(modified_hsmm$sojourn$M$d)
    modified_hsmm$sojourn$lE$d = dnorm(1:M, 
                                       mean = s_users$mean_lE_sojourn, 
                                       sd = s_users$sd_lE_sojourn)
    modified_hsmm$sojourn$Lut$d = dnorm(1:M, 
                                        mean = s_users$mean_Lut_sojourn, 
                                        sd = s_users$sd_Lut_sojourn)
    
    # specifying modified model
    modified_hsmm = specify_hsmm(J = modified_hsmm$J,
                                 init = modified_hsmm$init, trans = modified_hsmm$transition,
                                 sojourn = modified_hsmm$sojourn, 
                                 marg_em_probs = modified_hsmm$marg_em_probs,
                                 censoring_probs = modified_hsmm$censoring_probs)
    
    # simulating sequence 
    Xsim_i = simulate_hsmm(model = modified_hsmm, seq_id = as.character(sid), n_state_transitions = 100)
  })


```

__Tracking behavior categories__

```{r datasets-Xsim2}

Xsim2  = Xsim %>% 
  mutate(transitions = rbinom(nrow(Xsim), size = 1, prob = 0.001)) %>% 
  group_by(seq_id) %>% 
  mutate(tracking_sequence = 1+cumsum(transitions)) %>% 
  group_by(seq_id, tracking_sequence) %>% 
  mutate(tracking_category = sample(c("b","bp","btm","full"),1)) %>% 
  mutate(temp = ifelse(tracking_category %in% c("b","bp"), NA, temp),
         mucus = ifelse(tracking_category %in% c("b","bp"), NA, mucus),
         LH = ifelse(tracking_category %in% c("b","bp","btm"), NA, LH),
         preg = ifelse(tracking_category %in% c("b","btm"), NA, preg)) %>%
  ungroup() %>% 
  left_join(., s_users %>% mutate(seq_id = as.character(seq_id)), by = c("seq_id")) %>% 
  select(-transitions, -tracking_sequence)
  
write_feather(Xsim2, path = "../Data/synthetic_data/Xsim.feather")

```


__Visualization of a synthetic time-series example__


```{r datasets-synthetic-example,fig.width=12, fig.height=3, fig.cap="Example of a synthetic time-series."}

plot_hsmm_seq(X = Xsim2 %>%  filter(seq_id == 4), model = R_hsmm)

```


__Synthetic dataset characteristics__

```{r datasets-table-tracking-cat}

table(tracking_category = Xsim2$tracking_category)  %>% 
  kable(., format = "latex", booktabs = T, 
        caption = "Number of data-point in each tracking category.")

```


```{r datasets-histogram-time-series-length, fig.width=4, fig.height=2, fig.cap="Histogram of simulated sequence length."}

table(seq_id = Xsim2$seq_id) %>% 
  data.frame() %>% 
  set_colnames(c("seq_id","sequence_length")) %>% 
  ggplot(., aes(x = sequence_length)) +
  geom_histogram(bins = 50) + expand_limits(x = 0)

```


\newpage

## Kindara app data pre-processing and partial manual labeling. {#kindaraprep}


__pre-processing function__

```{r datasets-kindara-prepare-obs-fun}
source("Scripts/00_function_prepare_obs.R")
```

__pre-processing__

```{r datasets-kindara-preproc}

days = read_feather(path = paste0(IO$tmp_data, "days_selected_users.feather"))

X = purrr::map_dfr(
  .x = unique(days$user_id),
  .f = function(uid) prepare_obs(d = days %>% filter(user_id == uid))
) %>% 
  rename(seq_id = user_id, t = rel_date)

write_feather(X, path = paste0(IO$output_data, "processed_app_data.feather"))

```


__manual labeling__

```{r datasets-kindara-ML}

ML = read_feather(path = paste0(IO$output_data, "ML.feather"))

```


```{r datasets-kindara-ML-add, eval = FALSE}

load("../Data/models/R_hsmm.Rdata", verbose = TRUE)

ML = label_sequences(model = R_hsmm, X = X, ground_truth = ML)

write_feather(ML, path = paste0(IO$output_data, "ML.feather"))

```



