---
title: "Datasets"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---


```{r datasets-setup, include = FALSE, eval = TRUE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
source("Scripts/00_setup.R")
```

# Datasets


## Synthetic data generation


To generate synthetic data for $N$ fictive users, we first define some key characteristics of these users. First, their highest tracking frequency is defined via a parameter $\alpha$. Then, we randomly sample their menstrual characteristics such as the length and regularity of their follicular and luteal phases or the noise in their temperature measurements from realistic priors.

### Simulating realistic tracking behavior

Because we want to evaluate the performance of our framework given different level of missing data in a realistic context, we first define censoring probabilities for each state and variable that reflect the tracking behavior of a dedicated user whose purpose is to achieve or avoid pregnancy. A dedicated user would not track all the features every day but would instead systematically report the relevant features at the relevant moments of their cycle.
For example, they would report ovulation test in the days leading and following ovulation, not during their menses.

Once these censoring probabilities are defined, we can modulate them with a parameter $\alpha$ reflecting that some users are less dedicated than others.

We specify realistic censoring probabilities of a dedicated user as follow:

```{r datasets-dedicated-tracking}


load("../Data/models/R_hsmm.Rdata", verbose = TRUE)

p = R_hsmm$censoring_probs$p
p_ = p
p_[c("M")] = 0.01
p_[c("lE","Lut","Ano","AB","P","L","lEpL")] = 0.05
p_[c("hE","preO","O","postO")] = 0.025
p_[c("PL","PB1","PP")] = 0.9
p_[c("PB2","PB3","BF")] = 0.98
p_[c("B")] = 0.2

q = R_hsmm$censoring_probs$q
q_ = q
q_["LH", c("hE","preO","O","postO")] = c(0.3,0.1,0.1,0.1)
q_["mucus", c("hE","preO","O","postO")] = 0.1
q_["temp", c("hE","preO","O","postO","Ano")] = 0.1
q_["temp", c("Lut","P")] = 0.3
q_["temp", c("PL","PB1")] = 0.75

```


```{r datasets-p-expected-table}

p_ %>% t() %>% 
  kable(., 
        format = "latex", 
        booktab = TRUE, 
        caption = "Expected censoring probabilities p.") %>% 
  kable_styling(full_width = T, latex_options = c("hold_position")) 

```

```{r datasets-q-expected-table}

q_ %>% kable(., 
        format = "latex", 
        booktab = TRUE, 
        caption = "Expected marginal censoring probabilities q.") %>% 
  kable_styling(full_width = T, latex_options = c("hold_position")) 

```



We now modulate these probabilities with the parameter $\alpha$:

```{r datasets-missing-probs}

alpha_ref = c(0,0.8,1)
alphas = seq(0,1,by = 1/5)
alpha_levels = c("absent","rare","occasional","expected","diligent","persistent")

missing_probs = 
  map_dfr(
    .x = 1:R_hsmm$J,
    .f = function(state){
      modulated_p = spline(x = alpha_ref, y = c(1,p_[state],0), method = "hyman", xout = alphas)
      modulated_qs = 
        map_dfr(
          .x = R_hsmm$marg_em_probs %>% names(),
          .f = function(var){
            modulated_q = spline(x = alpha_ref, y = c(1,q_[var,state],0), method = "hyman", xout = alphas)
            data.frame(variable = var,
                       alpha = alphas,
                       q = modulated_q$y)
          }
        )
      missing_prob_this_state = 
        left_join(
          data.frame(state = state,
                     alpha_level = alpha_levels,
                     alpha = alphas,
                     p =  modulated_p$y
          ),
          modulated_qs,
          by = c("alpha")
        )
    }
  )

missing_probs = 
  missing_probs %>% 
  mutate(
    q = ifelse(variable == "bleeding", 0, q), # bleeding is never missing if the app is open
    missing_prob = p + (1-p)*q,
    alpha_level = alpha_level %>% factor(., levels = alpha_levels),
    state_name = R_hsmm$state_names[state] %>% factor(., levels = R_hsmm$state_names),
    state_color = R_hsmm$state_colors[state])

```


```{r datasets-expected-missingness, echo = FALSE, results='hide', fig.height=9, fig.cap="Probabilities for a variable to be reported given different tracking behaviors."}

ggplot(missing_probs,
       aes(x = alpha_level, y = 1-missing_prob, fill = state_color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity() +
  ylab("Probability of being reported") +
  xlab("Simulated tracking frequency") +
  facet_grid(state_name ~ variable) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        strip.text.y = element_text(angle = 0, hjust = 0))


```



__Synthetic users characteristics__

```{r datasets-synthetic-users}

N_per_alpha = 50 
N_users = (length(alpha_levels)-1) * N_per_alpha

s_users = data.frame(
  seq_id = 1:N_users, 
  alpha = rep(alphas[-1], N_per_alpha),
  alpha_level = rep(alpha_levels[-1],N_per_alpha),
  temp_sd = runif(N_users, min = 0.05, max = 0.3),
  mean_lE_sojourn = rnorm(N_users, mean = 10, sd = 3),
  sd_lE_sojourn = 0.2 + rpois(N_users, lambda = 2),
  mean_Lut_sojourn = rnorm(N_users, mean = 11, sd = 1),
  sd_Lut_sojourn = 0.2 + rpois(N_users, lambda = 0.75)
  )

```

__Generating time-series__


```{r datasets-Xsim, warning=FALSE}

# 139 sec = 2+ min for 64 time-series
Xsim = purrr::map_dfr(
  .x = s_users$seq_id,
  .f = function(sid){
    #cat("\n",sid, "\n")
    modified_hsmm = R_hsmm
    
    # modifying censoring probabilities
    this_user_alpha = s_users$alpha[s_users$seq_id == sid]
    modified_hsmm$censoring_probs$p = 
      missing_probs %>% filter(alpha == this_user_alpha) %>% 
      select(state, p) %>% arrange(state) %>% 
      distinct() %>% 
      select(p) %>% unlist() 
    modified_hsmm$censoring_probs$q = 
      missing_probs %>% filter(alpha == this_user_alpha) %>% 
      select(state, variable, q) %>% 
      pivot_wider(id_cols = variable, names_from = state, values_from = q) %>% 
      select(-variable) %>% as.matrix()
    
    # modifying emission distributions
    modified_hsmm$marg_em_probs$temp$params$sd = rep(s_users$temp_sd,R_hsmm$J)
    # modifying sojourns
    M = length(modified_hsmm$sojourn$M$d)
    modified_hsmm$sojourn$lE$d = dnorm(1:M, 
                                       mean = s_users$mean_lE_sojourn, 
                                       sd = s_users$sd_lE_sojourn)
    modified_hsmm$sojourn$Lut$d = dnorm(1:M, 
                                        mean = s_users$mean_Lut_sojourn, 
                                        sd = s_users$sd_Lut_sojourn)
    
    # specifying modified model
    modified_hsmm = specify_hsmm(J = modified_hsmm$J,
                                 init = modified_hsmm$init, trans = modified_hsmm$transition,
                                 sojourn = modified_hsmm$sojourn, 
                                 marg_em_probs = modified_hsmm$marg_em_probs,
                                 censoring_probs = modified_hsmm$censoring_probs)
    
    # simulating sequence 
    Xsim_i = simulate_hsmm(model = modified_hsmm, seq_id = as.character(sid), n_state_transitions = 100)
  })


```

__Altering the long-term tracking behavior of our fictive users__

```{r datasets-Xsim2}

Xsim2  = Xsim %>% 
  mutate(transitions = rbinom(nrow(Xsim), size = 1, prob = 0.001)) %>% 
  group_by(seq_id) %>% 
  mutate(tracking_sequence = 1+cumsum(transitions)) %>% 
  group_by(seq_id, tracking_sequence) %>% 
  mutate(tracking_category = sample(c("b","bp","btm","full"),1)) %>% 
  mutate(temp = ifelse(tracking_category %in% c("b","bp"), NA, temp),
         mucus = ifelse(tracking_category %in% c("b","bp"), NA, mucus),
         LH = ifelse(tracking_category %in% c("b","bp","btm"), NA, LH),
         preg = ifelse(tracking_category %in% c("b","btm"), NA, preg)) %>%
  ungroup() %>% 
  left_join(., s_users %>% mutate(seq_id = as.character(seq_id)), by = c("seq_id")) %>% 
  select(-transitions, -tracking_sequence)
  
write_feather(Xsim2, path = "../Data/synthetic_data/Xsim.feather")

```


__Visualization of synthetic time-series examples__

```{r datasets-selecting-examples-from-each-category}

examples = Xsim2 %>% 
  filter(alpha_level != "none") %>% 
  select(seq_id, tracking_category, alpha_level) %>% 
  group_by(seq_id, alpha_level) %>% 
  summarize(has_full_tracking = sum(tracking_category == "full") >= 100, .groups = "drop") %>% 
  filter(has_full_tracking) %>%
  group_by(alpha_level) %>% 
  slice_head(n = 1)
  

```


```{r datasets-synthetic-example,fig.width=12, fig.height=5, fig.cap="Examples of synthetic time-series."}

for(i in 1:nrow(examples)){
  ex_id = examples$seq_id[i]
  plot_hsmm_seq(X = Xsim2 %>%  filter(seq_id == ex_id), model = R_hsmm,
                title = str_c(examples$alpha_level[i], " tracking")) %>% 
    print()
}


```


__Synthetic dataset characteristics__

```{r datasets-table-tracking-cat}

table(tracking_category = Xsim2$tracking_category)  %>% 
  kable(., format = "latex", booktabs = T, 
        caption = "Number of data-point in each tracking category.") %>% 
  kable_styling(latex_options = c("hold_position")) 

```


```{r datasets-histogram-time-series-length, fig.width=4, fig.height=2, fig.cap="Histogram of simulated sequence length."}

table(seq_id = Xsim2$seq_id) %>% 
  data.frame() %>% 
  set_colnames(c("seq_id","sequence_length")) %>% 
  ggplot(., aes(x = sequence_length)) +
  geom_histogram(bins = 50) + expand_limits(x = 0)

```

```{r datasets-effective-tracking-frequency}

tracking_freq_table_synth_data = 
  Xsim2 %>% 
  filter(tracking_category == "full") %>% 
  group_by(alpha_level) %>% 
  mutate(any_log = !is.na(bleeding) | !is.na(LH) | !is.na(preg) | !is.na(mucus) | !is.na(temp),
         any_log = ifelse(!any_log,NA, 1)) %>% 
  summarize(across(c(bleeding, LH, preg, mucus, temp, any_log), .fns = function(x) sum(!is.na(x))),
            n = n(),
            .groups = "drop") %>% 
  mutate(Effective_tracking_frequency = (bleeding+LH+preg+mucus+temp)/5/n,
         Frequency_of_opening_the_app = any_log/n) %>% 
  select(alpha_level, Effective_tracking_frequency, Frequency_of_opening_the_app ) %>% 
  mutate(alpha_level = alpha_level %>% factor(., levels = alpha_levels)) %>% 
  arrange(alpha_level)

tracking_freq_table_synth_data %>% 
  rename(Tracking_category = alpha_level) %>% 
  kable(.,
        booktab = T,
        digits = 2,
        caption = "Effective tracking frequency (fraction of reported variables) and frequency of opening the app (at least one variable is logged at a given time-point) per simulated tracking behaviors.")

```




\newpage

## Kindara app data pre-processing and partial manual labeling. {#kindaraprep}


__pre-processing function__

```{r datasets-kindara-prepare-obs-fun}
source("Scripts/00_function_prepare_obs.R")
```

__pre-processing__

```{r datasets-kindara-preproc}

days = read_feather(path = paste0(IO$tmp_data, "days_selected_users.feather"))

X = purrr::map_dfr(
  .x = unique(days$user_id),
  .f = function(uid) prepare_obs(d = days %>% filter(user_id == uid))
) %>% 
  rename(seq_id = user_id, t = rel_date)

write_feather(X, path = paste0(IO$output_data, "processed_app_data.feather"))

```


__manual labeling__

```{r datasets-kindara-ML}

ML = read_feather(path = paste0(IO$output_data, "ML.feather"))

```


```{r datasets-kindara-ML-add, eval = FALSE}

load("../Data/models/R_hsmm.Rdata", verbose = TRUE)

ML = label_sequences(model = R_hsmm, X = X, ground_truth = ML)

write_feather(ML, path = paste0(IO$output_data, "ML.feather"))

```




```{r datasets-kindara-characteristics-and-tracking-frequency}

breaks_tracking_freq = c(0, tracking_freq_table_synth_data$Effective_tracking_frequency)
breaks_tracking_freq = (breaks_tracking_freq + lag(breaks_tracking_freq))/2
breaks_tracking_freq = c(0, breaks_tracking_freq[-1],1)


breaks_app_open = c(0, tracking_freq_table_synth_data$Frequency_of_opening_the_app)
breaks_app_open = (breaks_app_open + lag(breaks_app_open))/2
breaks_app_open = c(0, breaks_app_open[-1],1)

  
users_tracking_freq_cat = 
  X %>% 
  group_by(seq_id) %>% 
  mutate(any_log = !is.na(bleeding) | !is.na(LH) | !is.na(preg) | !is.na(mucus) | !is.na(temp),
         any_log = ifelse(!any_log,NA, 1)) %>% 
  summarize(across(c(bleeding, LH, preg, mucus, temp, any_log), .fns = function(x) sum(!is.na(x))),
            n = n(),
            .groups = "drop") %>% 
  mutate(Effective_tracking_frequency = (bleeding+LH+preg+mucus+temp)/5/n,
         Frequency_of_opening_the_app = any_log/n,
         tracking_freq_cat = Effective_tracking_frequency %>% 
           cut(breaks = breaks_tracking_freq,
               labels = c("none",tracking_freq_table_synth_data$alpha_level %>% as.character())),
         open_app_cat = Frequency_of_opening_the_app %>% 
           cut(breaks = breaks_app_open,
               labels = c("none",tracking_freq_table_synth_data$alpha_level %>% as.character()))
           ) 


users_tracking_freq_cat %>% 
  group_by(tracking_freq_cat) %>% 
  summarize(n = n(),
            perc = n/length(unique(X$seq_id)) * 100,
            .groups = "drop") %>% 
  kable(.,
        booktab = T,
        digits = 0,
        caption = "Number and percentage of users in our real-world dataset according to the tracking categories defined for our simulated data, based on the frequency with which they log features.")

```
```{r datasets-kindara-characteristics-and-tracking-frequency-2}


users_tracking_freq_cat %>% 
  group_by(open_app_cat) %>% 
  summarize(n = n(),
            perc = n/length(unique(X$seq_id)) * 100,
            .groups = "drop") %>% 
  kable(.,
        booktab = T,
        digits = 0,
        caption = "Number and percentage of users in our real-world dataset according to the tracking categories defined for our simulated data, based on the frequency with which they open the app.")

```


