---
title: "Predicting next period"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---


```{r predicting-setup, include = FALSE, eval = TRUE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
source("Scripts/00_setup.R")
```


# Predicting next period

In this section, we evaluate the ability of our model to learn the cycle characteristics of users and to predict the next period of irregular users (_i.e._ users with irregular cycle length).

To do so, we filter users (or part of their time-series) to keep those with irregular cycles and who log with a tracking frequency of at least 1/3 days. 

We compare the predictions from our model with the baseline prediction which is based on the mean or median cycle length of previous cycles.

## Loading models and decoding results

```{r predicting-loading-models, results="hide"}

load("../Data/models/R_hsmm.Rdata", verbose = TRUE)
load("../Data/models/M_hsmm.Rdata", verbose = TRUE)
load("../Data/models/T_hsmm.Rdata", verbose = TRUE)

```

```{r predicting-load-decodings}

RES.file = paste0(IO$output_data, "RES_app_data.feather")
RES = read_feather(path = RES.file)
rm(RES.file)

```

## Selecting users and cycles

```{r predicting-cycles-cycles-and-users-table}

RES = RES %>% 
  arrange(seq_id, t) %>% 
  group_by(seq_id) %>% 
  mutate(cycle_start = (state == 1) & ((lag(state) != 1) | (t == 1)),
         cycle_end = (state != 1) & (lead(state) == 1),
         cycle_number = cumsum(cycle_start)) %>% 
  group_by(seq_id, cycle_number) %>% 
  mutate(cycle_length = n()) %>% 
  ungroup()

cycles = RES %>% filter(cycle_start) %>% 
  select(-state, -cycle_start, -cycle_end)

users = cycles %>%
  group_by(seq_id) %>% 
  summarize(n_cycles = n(),
            median_cycle_length_all = median(cycle_length, na.rm = TRUE),
            median_cycle_length = median(cycle_length[cycle_length < 50], 
                                          na.rm = TRUE),
            sd_cycle_length_all = sd(cycle_length, na.rm = TRUE),
            sd_cycle_length = sd(cycle_length[cycle_length < 50], na.rm = TRUE),
            .groups = "drop")


```


```{r predicting-filtering-timeseries}

df = RES %>% 
  filter(
    # filter for tracking behavior that's not just bleeding and/or pregnancy tests
    str_detect(tracking_behavior, 
               str_c(c("b_tests","no_temp","no_mucus","full"), collapse = "|")),
    # filter for ovulatory cycles 
    state <= 7,
    # filter for ovulatory cycles in which the state prob is not too low (i.e. there was enough data to estimate ovulation states)
    prob > 0.3
  ) %>% 
  # we look for consecutive cycles fulfilling the criteria
  group_by(seq_id) %>% 
  mutate(new_cons_stretch = (t == min(t)) | ((t - lag(t)) > 1),
         cons_stretch_id = cumsum(new_cons_stretch)) %>% 
  # we make sure that cons_stretches start on a cycle start and end on a cycle end
  group_by(seq_id, cons_stretch_id) %>% 
  mutate(n_cycles = sum(cycle_start)) %>% 
  filter(n_cycles >= 5) %>% 
  mutate(t_first_cycle_start = min(t[cycle_start]),
         t_last_cycle_end = max(t[cycle_end])) %>% 
  ungroup() %>% 
  filter(t >= t_first_cycle_start, 
         t <= t_last_cycle_end) %>% 
  # we ensure that we have enough cycles in a stretch and that the cycle lengths show some variability
  group_by(seq_id, cons_stretch_id) %>% 
  mutate(n_cycles = length(unique(cycle_number)),
         cycle_length_diff = max(cycle_length) - min(cycle_length)) %>% 
  filter(
    # filter for at least 5 consecutive cycles
    n_cycles >= 5, 
    # filter for some irregularity in cycle length in this stretch
    cycle_length_diff >= 5) %>% 
  ungroup() 


df %>% select(seq_id, cons_stretch_id, cycle_number) %>% distinct() %>% nrow()
df %>% select(seq_id, cons_stretch_id) %>% distinct() %>% nrow()
df %>% select(seq_id) %>% distinct() %>% nrow()

```

## Predicting cycle length

### Building the input dataset

```{r predicting-load-time-series}

X = read_feather(path = paste0(IO$output_data, "processed_app_data.feather"))

```



```{r predicting-creating-input}

# for each user
# for each consecutive stretch
# for each consecutive 5-cycles group, 
# we keep the 4 first cycles (= training cycles)

input = data.frame()
input_baseline = data.frame()
output = data.frame()
fifth_cycle = data.frame()

for(u in unique(df$seq_id)){
  cat(u, "\t")
  dfu = df %>% filter(seq_id == u)
  for(s in unique(dfu$cons_stretch_id)){
    dfs = dfu %>% filter(cons_stretch_id == s)
    n_cycles = length(unique(dfs$cycle_number))
    first_cycle = min(dfs$cycle_number)
    for(cn in first_cycle:(first_cycle+n_cycles-5)){
      dfc = dfu %>% filter(cycle_number %in% cn:(cn+3))
      this_input = X %>%  
        filter(seq_id == u,
               t >= min(dfc$t),
               t <= max(dfc$t)+5) %>%  # we add 5 days so that the next period is included
        rename(o_seq_id = seq_id) %>% 
        mutate(seq_id = str_c(o_seq_id, "_",s,"_",cn))
      
      input = bind_rows(input, this_input)
      
      this_input_baseline = dfc %>% 
        select(seq_id, t, cycle_start, cycle_number, cycle_length) %>% 
        filter(cycle_start) %>% 
        rename(o_seq_id = seq_id) %>% 
        mutate(seq_id = str_c(o_seq_id, "_",s,"_",cn))
      
      input_baseline = bind_rows(input_baseline, this_input_baseline)
      
      dfc5 = dfu %>% filter(cycle_number %in% (cn+4))
      this_fifth_cycle = X %>% 
        filter(seq_id == u,
               t >= min(dfc5$t),
               t <= max(dfc5$t)) %>%
        rename(o_seq_id = seq_id) %>% 
        mutate(seq_id = str_c(o_seq_id, "_",s,"_",cn))
      
      fifth_cycle = bind_rows(fifth_cycle, this_fifth_cycle)

      this_output = dfu %>%  filter(cycle_number == cn+4) %>% 
        select(seq_id, t, cycle_number, cycle_length) %>%
        group_by(seq_id, cycle_number, cycle_length) %>% 
        summarize(t_cycle_end = max(t),
                  t_cycle_start = min(t),
                  .groups = "drop") %>% 
        rename(o_seq_id = seq_id) %>% 
        mutate(seq_id = str_c(o_seq_id, "_",s,"_",cn))
      output = bind_rows(output, this_output)
    }
  }
}

cat("\n")

```

### Training and predictions

```{r predicting-hsmm-model-for-ovulatory-cycles-only}

J = 7
marg_em_probs = R_hsmm$marg_em_probs
for(var in names(marg_em_probs)){
  if(marg_em_probs[[var]]$type == "non-par") marg_em_probs[[var]]$params$probs = marg_em_probs[[var]]$params$probs[,1:J]
  if(marg_em_probs[[var]]$type == "binom"){
    marg_em_probs[[var]]$params$size = marg_em_probs[[var]]$params$size[1:J]
    marg_em_probs[[var]]$params$prob = marg_em_probs[[var]]$params$prob[1:J]
  }
  if(marg_em_probs[[var]]$type == "norm"){
    marg_em_probs[[var]]$params$mean = marg_em_probs[[var]]$params$mean[1:J]
    marg_em_probs[[var]]$params$sd = marg_em_probs[[var]]$params$sd[1:J]
  }
}

# modify R_hsmm >> become P_hsmm (P for predictions)
P_hsmm = specify_hsmm(
  J = J ,
  state_names = R_hsmm$state_names[1:J],
  state_colors = R_hsmm$state_colors[1:J],
  init = R_hsmm$init[1:J],
  transition = R_hsmm$transition[1:J, 1:J],
  sojourn = R_hsmm$sojourn[1:J],
  marg_em_probs = marg_em_probs,
  censoring_probs = list(p = rep(0, J),
                         q = matrix(0.5, nrow = length(marg_em_probs), ncol = J))
)

```



```{r predicting-actual-predictions}


pred_file = "../Data/cycle_length_pred.feather"

if(!file.exists(pred_file)){
  
  # for each seq_id
  pred = purrr::map_dfr(
    .x = 1:length(unique(input$seq_id)),
    .f = function(i){
      cat(i, "\t")
      sid = unique(input$seq_id)[i]
      this_output = output %>% filter(seq_id == sid)
      
      # Baseline
      # The predicted cycle length is the mean of the cycle length of the 4 cycles. 
      # Uncertainty is the standard deviation over these 4 cycles.
      baseline_pred = input_baseline %>% filter(seq_id == sid) %>% 
        summarize(uncertainty = sd(cycle_length),
                  cycle_length = mean(cycle_length)) %>% 
        mutate(diff = cycle_length - this_output$cycle_length) %>% 
        full_join(., data.frame(cycleday = 1:this_output$cycle_length), 
                  by = character()) %>% 
        mutate(model = "Baseline",
               cycleday_backward = cycleday - this_output$cycle_length - 1)
      
      # HSMM
      # Train the model on the 4 cycles
      trained_model = fit_hsmm(model = P_hsmm, X = input %>% filter(seq_id == sid),
                               use_sojourn_prior = FALSE)
      # dec = predict_states_hsmm(model = trained_model$model, 
      #                           X = input %>% filter(seq_id == sid), 
      #                           method = "viterbi")
      # Predict from each day of the next cycle
      this_fifth_cycle = fifth_cycle %>% filter(seq_id == sid)
      t_cycle_start = min(this_fifth_cycle$t)
      hsmm_pred = purrr::map_dfr(
        .x = 1:this_output$cycle_length,
        .f = function(cd){
          # first we decode the 5th cycle until day cd
          dec = predict_states_hsmm(
            model = trained_model$model, 
            X = this_fifth_cycle %>% filter(t <= t_cycle_start + cd - 1), 
            ground_truth = data.frame(seq_id = sid, 
                                      t = t_cycle_start, 
                                      state = 1),
            trust_in_ground_truth = 1,
            method = "Viterbi")
          # we detect the last state transition
          last_state_transition = dec$state_seq %>% 
            group_by(state) %>% 
            summarize(t_last_transition = min(t), .groups = "drop") %>%
            arrange(-t_last_transition) %>% 
            slice_head()
          # we simulate X times and we compute the mean and sd of the predicted cycle lengths
          X = 100
          hsmm_pred_this_cd = purrr::map_dfr(
            .x = last_state_transition$state:7,
            .f = function(s){
              data.frame(state = s, i = 1:X, 
                         sojourn = sample(1:length(trained_model$model$sojourn[[s]]$d), 
                                          X, prob = trained_model$model$sojourn[[s]]$d,
                                          replace = TRUE))
            }
          ) %>% 
            group_by(i) %>% 
            summarize(cycle_length_from_last_transition = sum(sojourn), .groups = "drop") %>% 
            mutate(cycle_length = cycle_length_from_last_transition + last_state_transition$t_last_transition - t_cycle_start) %>% 
            summarize(uncertainty = sd(cycle_length),
                      cycle_length = mean(cycle_length)) %>% 
            mutate(cycleday = cd, diff = cycle_length - this_output$cycle_length)
          
          hsmm_pred_this_cd
        }
      ) %>% 
        mutate(model = "HSMM",
               cycleday_backward = cycleday - this_output$cycle_length - 1)
      
      
      this_seq_pred = bind_rows(baseline_pred, hsmm_pred) %>% 
        mutate(seq_id = sid)
      this_seq_pred
    }
  )
  
  write_feather(pred, path = pred_file)
  
}else{
  pred = read_feather(path = pred_file)
}


# ggplot(this_seq_pred, aes(x = cycleday_backward, y = diff, col = model)) +
#   geom_hline(yintercept = 0, col = "black")+
#   geom_line() +
#   geom_ribbon(aes(ymin = diff - uncertainty, ymax = diff + uncertainty, fill = model), col = NA, alpha = 0.5)

```

## Results

```{r predicting-results-HSMM, fig.cap="Distribution of the errors on the cycle length prediction from the HSMM simulations as one progresses through the cycles."}

# ggplot(pred %>% filter(model == "HSMM", cycleday_backward >= -25), aes(x = cycleday_backward, y = diff, col = model, group = seq_id))+
#   geom_line(alpha = 0.1) + geom_hline(yintercept = 0, linetype = 2)

ggplot(pred %>% 
         filter(model == "HSMM", cycleday_backward >= -25) %>% 
         mutate(diff_bin = round(diff)) %>% 
         group_by(cycleday_backward, diff_bin, model) %>% 
         summarize(n = n(), .groups = "drop") , 
       aes(x = cycleday_backward, y = diff_bin, fill = n))+
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  expand_limits(fill = 0) +
  ylim(c(-12,12)) + geom_hline(yintercept = 0, linetype = 2)

  
```

```{r predicting-results-comparison, fig.cap="Comparison of the distributions of the error on the cycle length prediction between the HSMM and the baseline method."}

ggplot(pred %>% filter(cycleday_backward %in% c(-25, -10)),
       aes(x = diff, fill = model)) +
  geom_density(alpha = 0.3, col = NA)+
  facet_grid(cycleday_backward ~ .)

```



```{r predicting-MSE-table}

pred_cycleday = c("At cycle start", "After ovulation\n(10 days before next cycle)")
pred_cycleday = factor(pred_cycleday, levels = pred_cycleday)

pred_res_table = pred %>% 
  filter(cycleday_backward == -10 | cycleday == 1) %>% 
  mutate(pred_cycleday = pred_cycleday[(cycleday != 1)+1]) %>% 
  group_by(pred_cycleday, model) %>% 
  summarize(MSE = mean(diff^2), .groups = "drop") %>% 
  pivot_wider(names_from = model, values_from = MSE, names_prefix = "MSE ") %>% 
  rename(`Prediction day` = pred_cycleday)

save(pred_res_table, file = "../Data/cycle_length_pred_summary_table.Rdata")

  kable(pred_res_table, digits = 2, caption = "MSE on the cycle length prediction for both model when made at the cycle start or 10 days before the next period.")

```



