---
title: "User selection for decoding and fitting"
author: "Laura Symul"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: true
    toc_depth: 5
    number_sections: true
    fig_caption: true
---

```{r idecod knitr setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r idecod setup, include = FALSE, eval = TRUE, cache = FALSE}
source("Scripts/00_setup.R")
```




## Users subset selection

To fit the model, we will use the time-series of a diverse subset of users.

We will select users from these different categories:
  
  1. Users with pregnancies

2. Users with many observations (mucus, temperature, LH tests)

3. Users with few observations (just period)

4. Users with very long time-series

5. Users with long and with short cycles

6. Both users who declared wanted to achieve and to avoid pregnancies


```{r idecod users selection}

users = read_feather(path = str_c(IO$output_data,"users.feather"))

top_n = 30

u1 = users %>% dplyr::filter(n_pos_preg_test > 0,
                             n_days_obs >= 365,
                             (batch %% 10) == 0 ) %>% mutate(u_type = "1") %>% head(top_n)

u2 = users %>% dplyr::filter(n_days_obs >= 365,
                             n_days_obs < 365*2,
                             (n_temp/n_days_obs >= 1/2) | (n_mucus/n_days_obs >= 1/4) |  (n_LH_test/n_days_obs >= 1/10),
                             (batch %% 10) == 0)%>% mutate(u_type = "2") %>% head(top_n)


u3 = users %>% dplyr::filter(n_days_obs >= 365,
                             n_days_obs < 365*2,
                             (n_temp < 10) & (n_mucus < 5) &  (n_LH_test<3),
                             (batch %% 10) == 0)%>% mutate(u_type = "3") %>% head(top_n)


u4 = users %>% dplyr::filter(n_days_obs >= 5*365,
                             (batch %% 10) == 0)%>%  mutate(u_type = "4") %>% head(top_n)


u5 = users %>% dplyr::filter(round(average_cycle) %in% c(20,40),
                             round(average_luteal) %in% c(11:14),
                             n_days_obs >= 365,
                             n_days_obs < 365*2,
                             (batch %% 10) == 0)%>%  mutate(u_type = "5") %>% head(top_n)


u = rbind(u1,u2,u3,u4,u5)
u = ddply(u, .(user_id, batch), summarize,
          u_type = min(u_type))
u = u %>%  arrange(u_type, batch, user_id)

dim(u)

selected_users = users %>% dplyr::filter(user_id %in% u$user_id)

write_feather(selected_users, path = str_c(IO$tmp_data,"selected_users.feather"))


```

```{r idecod retrieving their observations}

days_selected_users = foreach(b = unique(u$batch), .combine = rbind) %do% {
  cat(b,"\n")
  days = read_feather(path = str_c(IO$output_data,"Days/days_",b,".feather"))
  days = days %>% dplyr::filter(user_id %in% u$user_id)
  days
}

write_feather(days_selected_users, path = str_c(IO$tmp_data,"days_selected_users.feather"))

```









